Introduction to Causal Inference Models

### 2017-08-02 msm_introduction.Rmd
figured out iptw to td model.

If you've been to Society for Epidemiologic Research or any epi methods conference, you've probably heard something about "causal inference".

You've also probably heard a lot of gibberish about g-estimation, g-computation, g-formula, g-something, inverse-probability-of-treatment weighting, marginal structural models, doubly-robust, targeted maximum likelihood, etc.

And like me, you probably have no idea what any of those are, what they are used for, or even how to use them. But they are always referred to as being used in "causal inference", so these fancy-pants models must be special and the only way to find causal associations.

This presentation provides a general overview of things I've learned about these methods. I'm still confused about a lot of this topic area (like how is the g-formula different from g-estimation?), but hopefully this will server as a primer.

@@

To make things easier, let's load some functions I've made for g-computation and IPTW.

>>>>

Causal inference methods are very popular (particularly with methods epidemiologists)

Terms associated with thesee methods include: DAGs, g-estimation, g-computation, g-formula, g-something, inverse-probability-of-treatment weighting, marginal structural models, doubly-robust estimation, targeted maximum likelihood, etc.

This presentation provides an introduction to two of the more common of these methods.

## What's the big deal about causal inference methods?

 -

@@

```{r iptw td}
MSModelR::iptw

# define propensity models (intercept only model without covariates)
pr_a0_mod <- glm(A0 ~ 1, data = td_sim_data, family="binomial")
# probability of being treated at time 1
pr_a0 <- predict(pr_a0_mod, type = "response")
# define propensity a1 model
pr_a1_mod <- glm(A1 ~ 1, data = td_sim_data, family = "binomial")
# probability of being treated at time 2
pr_a1 <- predict(pr_a1_mod, type = "response")
# probability of a1 given z1
pr_a1_z_mod <- glm(A1 ~ Z1, data = td_sim_data, family = "binomial")
# probability of a1 given z1
pr_a1_z <- predict(pr_a1_z_mod, type = "response")

# create weights in dataframe
td_sim_data <- td_sim_data %>% 
  mutate(wt = ifelse(A0 == 1 & A1 == 1, (pr_a0/pr_a0)*(pr_a1/pr_a1_z),
    ifelse(A0 == 1 & A1 == 0, (pr_a0/pr_a0)*((1-pr_a1)/(1-pr_a1_z)),
    ifelse(A0 == 0 & A1 == 1, ((1-pr_a0)/(1-pr_a0))*(pr_a1/pr_a1_z),
    ifelse(A0 == 0 & A1 == 0, ((1-pr_a0)/(1-pr_a0))*((1-pr_a1)/(1-pr_a1_z)),NA
    )))))

# table of summarized weights
wt_table <- td_sim_data %>% 
  group_by(strata) %>% 
  summarise(A0 = mean(A0), Z1 = mean(Z1), A1 = mean(A1), Y = mean(Y), n = n(),
            wt = mean(wt), pseudo_n = n*wt)

wt_table

# weighted model
iptw_mod <- glm(Y ~ A0 + A1 + A0*A1, data = td_sim_data, weights = wt,
                family = "gaussian")

summary(iptw_mod)

sum(iptw_mod$coefficients[names(iptw_mod$coefficients)=="A0" | 
  names(iptw_mod$coefficients)=="A1" | names(iptw_mod$coefficients)=="A0*A1"],1)

```
To make things easier, let's load some functions I've made for g-computation and IPTW.
